---
title: "Assignment 06"
author: "Zhen Liu"
date: "2022-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```
## Contunuous random variables and limit laws
### Simulating data with the uniform distribution


#### Q1 

$$p_U(x) =   \left\{
\begin{array}{ll}
      1 &x\in [0,1] \\
      0&otherwise\\
\end{array} 
\right.$$
$$P(X\in[a,b]) = \int_{a}^{b}p_U(x)dx = 1 *length ([a,b]\cap[0,1]) = b-a$$

#### Q2

```{r}
set.seed(0) 
n <- 1000 
sample_X <- data.frame(U=runif(n)) %>% 
  mutate(X=case_when(
    (0<=U)&(U<0.25)~3,
    (0.25<=U)&(U<0.5)~10, 
    (0.5<=U)&(U<=1)~0)) %>% 
  pull(X)
```

Because the sample follows the distributions of Gaussian random variable, which is also called distribution

#### Q3 

```{r}
sample_X_0310<-function(a,b,n){
n <- n
sample_X_0310 <- data.frame(U=runif(n)) %>%
  mutate(X=case_when( 
    (0<=U)&(U<a)~3,
    (a<=U)&(U<(a+b))~10,
    ((a+b)<=U)&(U<=1)~0))%>% 
  pull(X)
}

```

#### Q4

```{r}
sample_X_0310(0.5,0.1,10000)

```
```{r}
average<- mean(sample_X_0310(0.5,0.1,10000))
average
```

Based on assignment 5 section 2.1, the value of Expectation X is 3*0.5 + 0.1*10 =2.5

The law if large numbers tells us that the sample average converges towards the expectation, for sequences of independent and identically distributed random varaibles.

#### Q5

```{r}
var(sample_X_0310(0.5,0.1,10000))
```

Var(X) is: 9*0.5 +100*0.1-9*0.5*0.5-100*0.1*0.1-60*0.5*0.1 = 8.25


#### Q6
1.
```{r}

beta<-seq(0.01,0.9,length = 100)
beta
```
```{r}
sample_X_0310_new<-function(a,b){
n <- 100
sample_X_0310 <- data.frame(U=runif(n)) %>%
  mutate(X=case_when( (0<=U)&(U<a)~3,
          (a<=U)&(U<(a+b))~10,
          ((a+b)<=U)&(U<=1)~0))
return(sample_X_0310)
}

```

2. 
```{r}
data_frame_X<- data.frame(beta)

```
3. 
```{r}


```


