---
title: "Assignment 06"
author: "Zhen Liu"
date: "2022-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```
## Contunuous random variables and limit laws
### Simulating data with the uniform distribution


#### Q1 

$$p_U(x) =   \left\{
\begin{array}{ll}
      1 &x\in [0,1] \\
      0&otherwise\\
\end{array} 
\right.$$
$$P(X\in[a,b]) = \int_{a}^{b}p_U(x)dx = 1 *length ([a,b]\cap[0,1]) = b-a$$

#### Q2

```{r}
set.seed(0) 
n <- 1000 
sample_X <- data.frame(U=runif(n)) %>% 
  mutate(X=case_when(
    (0<=U)&(U<0.25)~3,
    (0.25<=U)&(U<0.5)~10, 
    (0.5<=U)&(U<=1)~0)) %>% 
  pull(X)
```

Because the sample follows the distributions of Gaussian random variable, which is also called distribution

#### Q3 

```{r}
sample_X_0310<-function(a,b,n){
n <- n
sample_X_0310 <- data.frame(U=runif(n)) %>%
  mutate(X=case_when( 
    (0<=U)&(U<a)~3,
    (a<=U)&(U<(a+b))~10,
    ((a+b)<=U)&(U<=1)~0))%>% 
  pull(X)
}

```

#### Q4

```{r}
sample_X_0310(0.5,0.1,10000)

```
```{r}
average<- mean(sample_X_0310(0.5,0.1,10000))
average
```

Based on assignment 5 section 2.1, the value of Expectation X is 3*0.5 + 0.1*10 =2.5

The law if large numbers tells us that the sample average converges towards the expectation, for sequences of independent and identically distributed random varaibles.

#### Q5

```{r}
var(sample_X_0310(0.5,0.1,10000))
```

Var(X) is: 9*0.5 +100*0.1-9*0.5*0.5-100*0.1*0.1-60*0.5*0.1 = 8.25


#### Q6
1.
```{r}

beta<-seq(0.01,0.9,length = 100)

class(beta)
```

2.
```{r}
sample_X_0310_new<-function(b){
n <- 100
sample_X_0310 <- data.frame(U=runif(n)) %>%
  mutate(X=case_when( 
    (0<=U)&(U<0.1)~3,
    (0.1<=U)&(U<(0.1+b))~10,
    ((0.1+b)<=U)&(U<=1)~0))%>% 
  pull(X)
}


```


```{r}
Expectation<-function(b){
  return(mean(3*0.1 + b*10))
}

```

```{r}

data_frame_X<-data.frame(beta)

```
```{r}
data_frame_X%>%
  mutate(samplemean=map_dbl(.x=beta,.f=~mean(sample_X_0310_new(b=.x))))%>%
  mutate(Expectation=map_dbl(.x=beta,.f=~Expectation(b=.x)))
```



#### Q7

```{r}
  library(ggplot2)
```

```{r}

x<-data_frame_X$beta
y<-data_frame_X$samplemean
y_1<-data_frame_X$Expectation

plot(x,y,col="red")
lines(x,y_1,col="green")




```
### Exponential distribution

#### Q1

$$P_\lambda(x) = \int_{-\infty}^{\infty}p_\lambda(x)dx =  1-0 =1$$

#### Q2


```{r}
my_cdf_exp<-function(x,lambda){
  if(x<0){
  return(0)
  }
  else{
    return(1-10^-(lambda*x))
  }
}
```


```{r}
lambda <- 1/2 
map_dbl(.x=seq(-1,4), .f=~my_cdf_exp(x=.x,lambda=lambda) )
```
```{r}

test_inputs <- seq(-1,10,0.1) 
my_cdf_output <- map_dbl(.x=test_inputs, .f=~my_cdf_exp(x=.x,lambda=lambda)) 

```

#### Q3

```{r}
my_quantile_exp<-function(x,lambda){

  y<-1/my_cdf_exp(x,lambda)
  
}
```

```{r}

test_inputs <- seq(0.01,0.99,0.01) 
my_cdf_output <- map_dbl(.x=test_inputs, .f=~my_quantile_exp(x=.x,lambda=lambda)) 

```

#### Q4

The mean of population is ?

### The Binomial distribution and the central limit theorem


#### Q1

$$
P(Z=z) = p^z*(1-p)^(1-z),z =0,1\\
E(Z) = \sum_{z = 0}^{1} z P(z) = \sum_{z = 0}^{1} z P^z(1-p)^(1-z) = 0+p = p$$
$$E(Z^2) = \sum_{z = 0}^{1} z^2 P(z) = \sum_{z = 0}^{1} z^2 p^z (1-p)^(1-z) = 0+p = p $$
$$Var(Z) = E(Z^2)-(E(Z))^2 = p-p^2$$

#### Q2

1.
```{r}

x <- seq(0,50,by =1)

pmf<- dbinom(x,size = 50,prob = 0.7)

binom_df<-data.frame(x,pmf)
```
2. 

```{r}
head(binom_df,3)
```


#### Q3


```{r}
x<-seq(0,50,by =0.01)

pdf<-dnorm(x,mean = 35,sd = 3.24)

gaussian_df<-data.frame(x,pdf)

head(gaussian_df)
```

#### Q4

```{r}

colors<-c("Gaussianpdf"="red","Binomialpmf"="blue") 
fill<-c("Gaussianpdf"="white","Binomialpmf"="white") 


ggplot()+labs(x="x",y="Probability")+theme_bw()+ 
  geom_line(data=gaussian_df,aes(x,y=pdf,color="Gaussianpdf"),size=2)+ 
  geom_col(data=binom_df,aes(x=x,y=pmf,color="Binomialpmf",fill="Binomialpmf"))+ 
  scale_color_manual(name="myLegend",values=colors)+ scale_fill_manual(name="myLegend",values=fill)+ xlim(c(20,50))

```
### The Guassian distribution

#### Q1

```{r}


x <- seq(-10, 10, by = .1)
y <- dnorm(x, mean=1, sd=1)
y_1<-dnorm(x, mean=1, sd=1.41)
y_2<-dnorm(x, mean=1, sd=1.73)
plot(x, y,type = 'l',col='red')
lines(x, y_1,col='green')
lines(x, y_2,col='blue')

```

#### Q2

```{r}
x <- seq(-10, 10, by = .1)
y <- pnorm(x, mean=1, sd=1)
y_1<-pnorm(x, mean=1, sd=1.41)
y_2<-pnorm(x, mean=1, sd=1.73)
plot(x, y,type = 'l',col='red')
lines(x, y_1,col='green')
lines(x, y_2,col='blue')
```


#### Q3


```{r}
x <- seq(-10, 10, by = .1)
y <- qnorm(dnorm(x, mean=1, sd=1))
y_1<-qnorm(dnorm(x, mean=1, sd=1.41))
y_2<-qnorm(dnorm(x, mean=1, sd=1.73))
plot(x, y,type = 'l',col='red')
lines(x, y_1,col='green')
lines(x, y_2,col='blue')


```
reciprocal



#### Q4


```{r}
set.seed(0)
standardGaussianSample<-c(rnorm(100, mean = 0,sd=1))

```


#### Q5

```{r}



```

#### Q6


```{r}

```

#### Q7

```{r}

```


## Location estimators with Gaussian data

```{r}
set.seed(0) 
num_trials_per_sample_size <- 1000 
min_sample_size <- 30 
max_sample_size <- 500 
sample_size_inc <- 5 
mu_0 <- 1 
sigma_0 <- 3 


simulation_df<-crossing(trial=seq(num_trials_per_sample_size), 
                        sample_size=seq(min_sample_size, 
                                        max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),
                         .f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
  mutate(sample_md=map_dbl(.x=simulation,.f=median)) %>%
  group_by(sample_size)%>%
  summarise(msq_error_md=mean((sample_md-mu_0)^2))



```

#### Q1


```{r}
median(simulation_df$msq_error_md)
```

#### Q2

```{r}
simulation_df<-crossing(trial=seq(num_trials_per_sample_size), 
                        sample_size=seq(min_sample_size, 
                                        max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),
                         .f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
  mutate(sample_md=map_dbl(.x=simulation,.f=median)) %>%
  group_by(sample_size)%>%
  summarise(msq_error_md=mean((sample_md-mu_0)^2))
 



```


```{r}

simulation_df_mean<-crossing(trial=seq(num_trials_per_sample_size), 
                        sample_size=seq(min_sample_size, 
                                        max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),
                         .f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
  mutate(sample_mn=map_dbl(.x=simulation,.f=mean)) %>%
  group_by(sample_size)%>%
  summarise(msq_error_mn=mean((sample_mn-mu_0)^2))



```

```{r}
x <- simulation_df$sample_size
y <- simulation_df$msq_error_md
y_1<- simulation_df_mean$msq_error_mn
plot(x, y,type = 'l',col='red')
lines(x,y_1,col ='green')
```

