---
title: "Assignment 06"
author: "Zhen Liu"
date: "2022-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```
## Contunuous random variables and limit laws
### Simulating data with the uniform distribution


#### Q1 

$$p_U(x) =   \left\{
\begin{array}{ll}
      1 &x\in [0,1] \\
      0&otherwise\\
\end{array} 
\right.$$
$$P(X\in[a,b]) = \int_{a}^{b}p_U(x)dx = 1 *length ([a,b]\cap[0,1]) = b-a$$

#### Q2

```{r}
set.seed(0) 
n <- 1000 
sample_X <- data.frame(U=runif(n)) %>% 
  mutate(X=case_when(
    (0<=U)&(U<0.25)~3,
    (0.25<=U)&(U<0.5)~10, 
    (0.5<=U)&(U<=1)~0)) %>% 
  pull(X)
```

Because the sample follows the distributions of Gaussian random variable, which is also called distribution

#### Q3 

```{r}
sample_X_0310<-function(alpha,beta,n){
sample_X <- data.frame(U=runif(n)) %>%
  mutate(X=case_when( 
    (0<=U)&(U<alpha)~3,
    (alpha<=U)&(U<(alpha+beta))~10,
    ((alpha+beta)<=U)&(U<=1)~0))%>% 
  pull(X)
  return(sample_X)
}

```

#### Q4

```{r}
# answer
n<-10000
alpha<- 1/2
beta<- 1/10
sample_X<-sample_X_0310(alpha,beta,n)
mean(sample_X)

```
```{r}
average<- mean(sample_X_0310(0.5,0.1,10000))
average
```

Based on assignment 5 section 2.1, the value of Expectation X is 3*0.5 + 0.1*10 =2.5

The law if large numbers tells us that the sample average converges towards the expectation, for sequences of independent and identically distributed random varaibles.

#### Q5
```{r}
#answer
var(sample_X)
```

```{r}
var(sample_X_0310(0.5,0.1,10000))
```

Var(X) is: 9*0.5 +100*0.1-9*0.5*0.5-100*0.1*0.1-60*0.5*0.1 = 8.25


#### Q6
1.
```{r}

beta<-seq(0.01,0.9,length = 100)

class(beta)
```

2.
```{r}
sample_X_0310_new<-function(b){
n <- 100
sample_X_0310 <- data.frame(U=runif(n)) %>%
  mutate(X=case_when( 
    (0<=U)&(U<0.1)~3,
    (0.1<=U)&(U<(0.1+b))~10,
    ((0.1+b)<=U)&(U<=1)~0))%>% 
  pull(X)
}


```


```{r}
Expectation<-function(b){
  return(mean(3*0.1 + b*10))
}

```

```{r}

data_frame_X<-data.frame(beta)

```
```{r}
data_frame_X%>%
  mutate(samplemean=map_dbl(.x=beta,.f=~mean(sample_X_0310_new(b=.x))))%>%
  mutate(Expectation=map_dbl(.x=beta,.f=~Expectation(b=.x)))
```

```{r}
#### Q7 answer
n=100 
alpha=1/10 
samples<-data.frame(beta=seq(0,9/10,0.01))%>%
  mutate(sample_X=map(beta,~sample_X_0310(alpha,.x,n)))%>%
  mutate(samplemean=map_dbl(sample_X,mean))%>%
  mutate(Expectation=3*alpha+10*beta)
```


```{r}
  library(ggplot2)
```

```{r}

x<-data_frame_X$beta
y<-data_frame_X$samplemean
y_1<-data_frame_X$Expectation

plot(x,y,col="red")
lines(x,y_1,col="green")




```


```{r}
#answer

df<-samples%>%
  pivot_longer(cols=c(samplemean,Expectation), names_to='name',values_to='value')

ggplot(df,aes(x=beta,y=value,color=name))+ geom_point()+theme_bw()
```



### Exponential distribution

#### Q1

$$P_\lambda(x) = \int_{-\infty}^{\infty}p_\lambda(x)dx =  1-0 =1$$

#### Q2


```{r}
my_cdf_exp<-function(x,lambda){
  if(x<0){
  return(0)
  }

    return(1-exp(-lambda*x))
}
```


```{r}
lambda <- 1/2 
map_dbl(.x=seq(-1,4), .f=~my_cdf_exp(x=.x,lambda=lambda) )
```
```{r}

test_inputs <- seq(-1,10,0.1) 
my_cdf_output <- map_dbl(.x=test_inputs, .f=~my_cdf_exp(x=.x,lambda=lambda)) 
inbuilt_cdf_output <- map_dbl(.x=test_inputs,.f=~pexp(q=.x,rate=lambda)) 
all.equal(my_cdf_output,inbuilt_cdf_output)

```

#### Q3

```{r}
my_quantile_exp<-function(x,lambda){
  if (p<=0) return(-Inf)
  return(log(1-p)/(-lambda))
  
}
```

```{r}
test_inputs <- seq(0.01, 0.99, 0.01) 

my_quantile_output <- map_dbl(.x=test_inputs, .f=~my_quantile_exp(p=.x,lambda=lambda)) 

inbuilt_quantile_output <- map_dbl(.x=test_inputs,.f=~qexp(p=.x,rate=lambda))

all.equal(my_quantile_output,inbuilt_quantile_output)

```

#### Q4

The mean of population is ?

### The Binomial distribution and the central limit theorem


#### Q1

$$
P(Z=z) = p^z*(1-p)^(1-z),z =0,1\\
E(Z) = \sum_{z = 0}^{1} z P(z) = \sum_{z = 0}^{1} z P^z(1-p)^(1-z) = 0+p = p$$
$$E(Z^2) = \sum_{z = 0}^{1} z^2 P(z) = \sum_{z = 0}^{1} z^2 p^z (1-p)^(1-z) = 0+p = np $$
$$Var(Z) = E(Z^2)-(E(Z))^2 = n(p-p^2)$$

#### Q2

1.
```{r}

x <- seq(0,50,by =1)

pmf<- dbinom(x,size = 50,prob = 0.7)

binom_df<-data.frame(x,pmf)
```
2. 

```{r}
head(binom_df,3)
```


#### Q3


```{r}
x<-seq(0,50,by =0.01)

pdf<-dnorm(x,mean = 35,sd = 3.24)

gaussian_df<-data.frame(x,pdf)

head(gaussian_df)
```

#### Q4

```{r}

colors<-c("Gaussianpdf"="red","Binomialpmf"="blue") 
fill<-c("Gaussianpdf"="white","Binomialpmf"="white") 


ggplot()+labs(x="x",y="Probability")+theme_bw()+ 
  geom_line(data=gaussian_df,aes(x,y=pdf,color="Gaussianpdf"),size=2)+ 
  geom_col(data=binom_df,aes(x=x,y=pmf,color="Binomialpmf",fill="Binomialpmf"))+ 
  scale_color_manual(name="myLegend",values=colors)+ scale_fill_manual(name="myLegend",values=fill)+ xlim(c(20,50))

```
### The Guassian distribution

#### Q1

```{r}


x <- seq(-10, 10, by = .1)
y <- dnorm(x, mean=1, sd=1)
y_1<-dnorm(x, mean=1, sd=1.41)
y_2<-dnorm(x, mean=1, sd=1.73)
plot(x, y,type = 'l',col='red')
lines(x, y_1,col='green')
lines(x, y_2,col='blue')

```

```{r}

f_musigma <-function(mu, sigma, x){ 
  y = (1/sigma/sqrt(2*pi)) * exp(-0.5*((x-mu)/sigma)^2)
  return (y) 
  } 
df_density <- data.frame( x = seq(-4, 6, 0.1) )%>%
  mutate( '1' = f_musigma(mu=1,sigma=1,x=x), '2' = f_musigma(mu=1,sigma=sqrt(2),x=x), '3' = f_musigma(mu=1,sigma=sqrt(3),x=x) ) 

df_density_longer = pivot_longer(df_density, col = c('1','2','3'), names_to = 'Variance', values_to = 'Density')

ggplot(df_density_longer, aes(x=x, y = Density, linetype=Variance, color=Variance)) + geom_line() + theme_bw()


```






#### Q2

```{r}
x <- seq(-10, 10, by = .1)
y <- pnorm(x, mean=1, sd=1)
y_1<-pnorm(x, mean=1, sd=1.41)
y_2<-pnorm(x, mean=1, sd=1.73)
plot(x, y,type = 'l',col='red')
lines(x, y_1,col='green')
lines(x, y_2,col='blue')
```
```{r}
# Answer
df<-data.frame(x=seq(-4,6,0.1))%>%
  mutate('1'=pnorm(mean=1,sd=1,q=x), '2'=pnorm(mean=1,sd=sqrt(2),q=x), '3'=pnorm(mean=1,sd=sqrt(3),q=x))

df_longer=pivot_longer(df,col=c('1','2','3'), names_to='Variance',values_to='CDF')

ggplot(df_longer,aes(x=x,y=CDF,linetype=Variance,color=Variance))+ geom_line()+theme_bw()+ylab('Cumulativedistributionfunction')

```


#### Q3


```{r}
x <- seq(0, 1, by = 0.005)
y <- qnorm(dnorm(x, mean=1, sd=1))
y_1<-qnorm(dnorm(x, mean=1, sd=1.41))
y_2<-qnorm(dnorm(x, mean=1, sd=1.73))
plot(x, y,type = 'l',col='red')
lines(x, y_1,col='green')
lines(x, y_2,col='blue')


```
reciprocal
```{r}
#answer

df<-data.frame(x=seq(0,1,0.005))%>%
  mutate('1'=qnorm(mean=1,sd=1,p=x), 
         '2'=qnorm(mean=1,sd=sqrt(2),p=x), 
         '3'=qnorm(mean=1,sd=sqrt(3),p=x)) 

df_longer=pivot_longer(df,col=c('1','2','3'), names_to='Variance',values_to='Quantile')

ggplot(df_longer,aes(x=x,y=Quantile,linetype=Variance,color=Variance))+ 
  geom_line()+theme_bw()
```



#### Q4


```{r}
set.seed(0)
standardGaussianSample<-rnorm(100, mean = 0,sd=1)
head(standardGaussianSample)
```


#### Q5

```{r}

mean1Var3GaussianSampleA= standardGaussianSample*sqrt(3) + 1

head(mean1Var3GaussianSampleA)

```

#### Q6


```{r}
set.seed(0) 
mean1Var3GaussianSampleB = rnorm(100, mean=1, sd=sqrt(3)) 
head(mean1Var3GaussianSampleB)
```

```{r}
all.equal(mean1Var3GaussianSampleA, mean1Var3GaussianSampleB)
```



#### Q7

```{r}

colors<-c("Population density"="red", "Sample kernel density"="blue",
          "Population mean"="green", "Sample mean"="violet") 
linetypes<-c("Population density"="solid", "Sample kernel density"="dashed", "Population mean"="solid", "Sample mean"="dashed") 

ggplot() + labs(x="x",y="Density") + theme_bw() + 
  # 1. create plot of theoretical density 
  geom_line(data=select(df_density, x, "3"), aes(x=x,y=`3`,color="Population density", linetype="Population density")) + 
  # 2. add in kernel density plot from real sample 
  geom_density(data=data.frame(x=mean1Var3GaussianSampleA), aes(x=x,color="Sample kernel density", linetype="Sample kernel density")) + 
  # 3. vertical lines 
  geom_vline(aes(xintercept=1,color="Population mean", linetype="Population mean")) + geom_vline(aes(xintercept=mean(mean1Var3GaussianSampleA),color="Sample mean",linetype="Sample mean")) +
  
  scale_color_manual(name = "Legend", values=colors) + scale_linetype_manual(name="Legend", values=linetypes)


```


## Location estimators with Gaussian data

```{r}
set.seed(0) 
num_trials_per_sample_size <- 1000 
min_sample_size <- 30 
max_sample_size <- 500 
sample_size_inc <- 5 
mu_0 <- 1 
sigma_0 <- 3 


simulation_df<-crossing(trial=seq(num_trials_per_sample_size), 
                        sample_size=seq(min_sample_size, 
                                        max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),
                         .f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
  mutate(sample_md=map_dbl(.x=simulation,.f=median)) %>%
  group_by(sample_size)%>%
  summarise(msq_error_md=mean((sample_md-mu_0)^2))



```

#### Q1


```{r}
median(simulation_df$msq_error_md)
```

#### Q2

```{r}
simulation_df<-crossing(trial=seq(num_trials_per_sample_size), 
                        sample_size=seq(min_sample_size, 
                                        max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),
                         .f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
  mutate(sample_md=map_dbl(.x=simulation,.f=median)) %>%
  group_by(sample_size)%>%
  summarise(msq_error_md=mean((sample_md-mu_0)^2))
 



```


```{r}

simulation_df_mean<-crossing(trial=seq(num_trials_per_sample_size), 
                        sample_size=seq(min_sample_size, 
                                        max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),
                         .f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
  mutate(sample_mn=map_dbl(.x=simulation,.f=mean)) %>%
  group_by(sample_size)%>%
  summarise(msq_error_mn=mean((sample_mn-mu_0)^2))



```

```{r}
x <- simulation_df$sample_size
y <- simulation_df$msq_error_md
y_1<- simulation_df_mean$msq_error_mn
plot(x, y,type = 'l',col='red')
lines(x,y_1,col ='green')
```
```{r}
#answer
set.seed(0) 
num_trials_per_sample_size<-1000 
min_sample_size<-30 
max_sample_size<-500 
sample_size_inc<-5 
mu_0<-1 
sigma_0<-3 

#createdataframeofallpairsofsample_sizeandtrial

simulation_df<-crossing(trial=seq(num_trials_per_sample_size), sample_size=seq(min_sample_size, max_sample_size,sample_size_inc))%>%
  #simulatesequencesofGaussianrandomvariables
  
  mutate(simulation=pmap(.l=list(trial,sample_size), .f=~rnorm(.y,mean=mu_0,sd=sigma_0)))%>%
  
  #computethesamplemedians
  
  mutate(sample_md=map_dbl(.x=simulation,.f=median))%>% mutate(sample_mn=map_dbl(.x=simulation,.f=mean))%>%
  group_by(sample_size) %>%
  
  summarise(msq_error_md=mean((sample_md-mu_0)^2),msq_error_mn=mean((sample_mn-mu_0)^2))


```


```{r}

simulation_df%>%
  pivot_longer(cols=c(msq_error_md,msq_error_mn), names_to="Estimator",values_to="msq_error")%>%
  mutate(Estimator=case_when(Estimator=="msq_error_md"~"Median", Estimator=="msq_error_mn"~"Mean"))%>%
  
  ggplot(aes(x=sample_size,y=msq_error,color=Estimator,linetype=Estimator))+ geom_smooth()+theme_bw()+xlab("Samplesize")+ylab("Meansquareerror")


```

