---
title: "Assignment04"
author: "Zhen Liu"
date: "2022-10-23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.Probability theory 

### 1.1 Rules of probability 

#### Q1
P({a}) = 0.5
P({b}) = 0.1
P({c})=0.4

#### Q2 

rule 1: P(∅) = 0, since q ∈ [0,1], hence P({0}) = 1−q ≥ 0, P({1}) = q ≥ 0, P({0,1}) = 1 satisfy that P(A) ≥ 0 for any event A ∈ E
rule 2: Ω = {0,1} and P({0,1}) = 1 which satisfy P(Ω) = 1 for sample space Ω
rule 3: P({0}) = 1−q, P({1}) = q, P({0,1}) = 1 = P({0}) + P({1}) which satisfy rule 3.

### 1.2 Deriving new properties from the rules of probability

#### Q1 Union of a finite sequence of disjoint events

Based on lecture 10 rule 3 proof,

a finite probability sapce is triple (Ω,E,p) and E = {A ⊆ Ω}
Ω is finite, $$P(∪^\infty_{i=1}A_i) =\sum_{i = 1}^{\infty}(\sum_{i = 1}^{n}p_i\mathbf {1}A_i(w_i)) $$
hence $$P(∪^n_{i=1}A_i) =\sum_{i = 1}^{n}p_i\mathbf {1}A_i(w_i) =  \sum_{i = 1}^{n}\mathbf{p}(A_i)$$

#### Q2 Probability of a complement 

Based on rule 2 and rule 3

$$P(S^c) + P(S)= 1.$$ hence $$P(S^c) = 1 - P(S)$$

#### Q3 The union bound 

P(A∪B)=P(A)+P(B)−P(A∩B)≤P(A)+P(B)
$$\text{It is possible that} \ P(S_{i-1} ∩ S_{i-2}) >0 \ \text{hence} P(∪^\infty_{i=1}S_i) < \sum_{i=1}^{\infty}P(S_i)$$

#### Q4

$$A∪B = (A ∩ B) ∪ (A∩ B^c) ∪ (A^c∩ B)$$
$$P(A)= P(A ∩ B)+P(A∩ B^c)$$
$$P(B)= P(A ∩ B)+P(B∩ A^c)$$
Based on rule 3 
$$P(A ∪ B) = P[(A ∩ B) ∪ (A∩ B^c) ∪ (A^c∩ B)]=
P(A ∩ B) + P(A∩ B^c) + P(A^c∩ B) = P(A) + P(B) - P(A ∩ B)$$

## 2. Finite probability spaces

### 2.1 Sampling with replacement

#### Q1

\begin{equation} 
 P(A_{z,22})=\binom{22}{z}*(0.3)^z*(0.7)^{22-z}


\end{equation}

#### Q2

```{r}

prob_red_spheres<-function(z){
  pro<- choose(22,z)*(0.3^z)*(0.7^(22-z))
  return(pro)
}

```
```{r}
prob_red_spheres(10)

```


#### Q3 

```{r}
library(tidyverse)
num_reds<- c(seq(22))
prob<-c(prob_red_spheres(num_reds))


prob_by_num_reds<- data.frame(num_reds,prob)

prob_by_num_reds %>% head(3)
```
#### Q4

```{r}
library(ggplot2)
ggplot(prob_by_num_reds,aes(x=num_reds,y=prob)) + geom_line()
```

#### Q5

```{r}
sample(10, 22, replace=TRUE)
```
```{r}
## Setting the random seed just once
set.seed(0)
for(i in 1:5){
print(sample(100,5,replace=FALSE))
# The result may well differ every time
}
## Resetting the random seed every time
```

```{r}
for(i in 1:5){
set.seed(1)
print(sample(100,5,replace=FALSE))
# The result should not change
}
```
```{r}

num_trials<-1000 # set the number of trials
set.seed(0) # set the random seed
sampling_with_replacement_simulation<-data.frame(trial=1:num_trials) %>%
mutate(sample_balls = map(.x=trial, ~sample(10,22, replace = TRUE)))
# generate collection of num_trials simulations

```

```{r}
sampling_with_replacement_simulation<-sampling_with_replacement_simulation%>%
  mutate(num_reds = map_dbl(.x=sample_balls,~sum(.x<=3)))

```

#### Q6


```{r}

num_reds_in_simulation<-sampling_with_replacement_simulation %>%
pull(num_reds)
# we extract a vector corresponding to the number of reds in each trial
prob_by_num_reds<-prob_by_num_reds %>%
mutate(predicted_prob=map_dbl(.x=num_reds,~sum(num_reds_in_simulation==.x))/num_trials)
# add a column which gives the number of trials with a given number of reds



```
#### Q7

```{r}
prob_by_num_reds %>%
rename(TheoreticalProbability=prob, EstimatedProbability=predicted_prob) %>%
pivot_longer(cols=c("EstimatedProbability","TheoreticalProbability"),
names_to="Type",values_to="count") %>%
ggplot(aes(num_reds,count)) +
geom_line(aes(linetype=Type, color=Type)) + geom_point(aes(color=Type)) +scale_linetype_manual(values = c("solid", "dashed"))+
theme_bw() + xlab("Number of reds") + ylab("Probabilities")


```

### 2.2 Sampling without replacement

#### Q1

1. First set a random seed;
```{r}
set.seed(0)


```
2. 
```{r}
num_trials_no<-1000
sample(100,10,replace = FALSE)
```

3. 
```{r}

sampling_without_replacement_simulation<-data.frame(trial=1:num_trials_no) %>%
mutate(sample_collections = map(.x=trial, ~sample(100,10, replace = FALSE)))

```
4. 
```{r}
sampling_without_replacement_simulation<-sampling_without_replacement_simulation%>%
  mutate(reds = map_dbl(.x=sample_collections,~sum(.x<=50)))%>%
  mutate(blues = map_dbl(.x=sample_collections,~sum((.x>50) *(.x<=80))))%>%
  mutate(greens = map_dbl(.x=sample_collections,~sum((.x>80))),
  )

  

```

5. 
```{r}
missing_values<-sampling_without_replacement_simulation%>%
  mutate(mun_in = pmin(reds,blues,greens))
```
6.
```{r}

portion=mean(missing_values$mun_in==0)

print(portion)
```




