---
title: "Assginment02"
author: "Zhen Liu"
date: "2022-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data Wrangling 
### 1.1 Select and Filter 
#### Q1
```{r load-packages, include=FALSE}

library(magrittr)
library(knitr)
library(tidyverse)
library(Stat2Data)
data("Hawks")
```


```{r}
hSF<-Hawks%>%
  filter(Species == 'RT'& Weight >= 1000)%>%
  select(Wing,Weight,Tail)
head(hSF,5)
```
#### Q2 
#### How many variables does the data frame hSF have? How mant examples does the data frame hSF have? How many observations? How many cases?

```{r}

hSF%>%nrow()
```

totally 3 variables Wing, Weight, Tail\
totally 398 obervations or cases or examples

### 1.2 The arrange function

#### Q1

```{r}
head(hSF %>%  arrange((Wing)),5)

```
### 1.3 Join and rename functions 

#### Q1

```{r}
Species <- c("CH","RT","SS")
species_name_full <- c("Cooper's","Red-tailed","Sharp-shinned")
hawkSpeciesNameCodes<- data.frame(Species, species_name_full)
```

#### Q2
```{r}
hawksFullName <- left_join(Hawks,hawkSpeciesNameCodes,by = "Species")

hawksFullName<- (hawksFullName%>%select(-Species)%>%rename(Species =species_name_full ))

```
#### Q3
```{r}
head(hawksFullName%>%select(Species,Wing,Weight),7)
```
### 1.4 The mutate function
#### Q1 
\[
\Bird-BMI: = 1000 × Weight/Wing-span^2.
\]

```{r}
hawksWithBMI<- Hawks %>%
  mutate(bird_BMI = 1000*(Weight/(Wing*Wing)) )%>%
  select(Species,bird_BMI)%>%
  arrange(desc(bird_BMI))

head(hawksWithBMI,8)
```

### 1.5 Summarize and group-by functions
#### Q1 


```{r}
hawksFullName%>%
  group_by(Species)%>%
  summarise(num_rows = n(), mn_wing= mean(Wing),md_wing = median(Wing), t_mn_wing = mean(Wing,trim = 0.1),b_wt_ratio= max(Wing/Tail))

```
#### Q2

```{r}

hawksFullName%>%
  group_by(Species)%>%
  select(Wing,Weight,Culmen,Hallux,Tail,StandardTail,Tarsus,Crop)%>%
  summarise(across(everything(),~sum(is.na(.x))))

```

## 2. Tidy data and iteration
### 2.1 Missing data and interation
#### Q1

```{r}
impute_by_mean<-function(x){ 
  mu<-mean(x,na.rm=1) #first compute the mean of x 
  impute_f<-function(z){ #coordinate-wise imputation
    if(is.na(z)){ 
      return(mu)#if z is na replace with mean 
    }else{ 
        return(z) #otherwise leave in place
    }
  } 
  return(map_dbl(x,impute_f)) #apply the map function to impute across vector
}

```

for tidyverse library 

#### Q2

```{r}
impute_by_median <- function(x){
  mu<-median(x,na.rm=1)
  impute_f<- function(z){
    if(is.na(z)){
      return(mu)
    }
    else{
      return(z)
    }
  }
return(map_dbl(x,impute_f))
}
```
```{r}
v<- c(1,2,NA,4)
impute_by_median(v)
```
#### Q3

```{r}
x<- c(seq(0,1,by =0.1 ))
y<-c(5*x+1)
df_xy<- data.frame(x,y)
df_xy%>%head(5)
```
```{r}

df_xy%>%
  mutate(z =map2_dbl(df_xy$x, df_xy$y, sum))



```

#### Q4
```{r}

sometimes_missing<-function(index,value){
  if (index%%5 ==0) {
    return(NA)
  }
 
  return(value)
}

sometimes_missing(15,25)

```
```{r}



df_xy_missing<- df_xy%>%
  mutate(y = map2_dbl(row_number(),df_xy$y,sometimes_missing))

df_xy_missing%>% head(10)

```
#### Q5

```{r}



df_xy_imputed<-df_xy_missing%>%
  mutate(y=impute_by_median(df_xy_missing$y))
  
df_xy_imputed

```
### 2.2 Tidying data with pivot functions


```{r}

library(readxl) # load the readxl library
folder_path<-"/Users/zhenliu15471/code/zliu15471/SCEM-Msc-datascience/SCEM-Msc-datascience/"  # set this to the name of the 
# directory containing "HockeyLeague.xlsx"
file_name<-"HockeyLeague.xlsx"  # set the file name
file_path<-paste(folder_path,file_name,sep="") 
# create the file_path
wins_data_frame<-read_excel(file_path,sheet="Wins") # read of a sheet from an xl file
```
```{r}
wins_data_frame %>% select(1:5)%>% head(3)
```
#### Q1

```{r}

wins_tidy<- wins_data_frame%>%
  pivot_longer(!1,names_to = 'Year',values_to = "value")%>%
  separate(value,into= c("Wins","Total"),sep="of")%>%
  rename("Team"=1)%>%
  mutate(Year = as.integer(Year),Wins = as.integer(Wins),Total = as.integer(Total))

wins_tidy%>%dim()


```


```{r}
wins_tidy%>%head(5)
```
#### Q2
```{r}
library(readxl) # load the readxl library
folder_path<-"/Users/zhenliu15471/code/zliu15471/SCEM-Msc-datascience/SCEM-Msc-datascience/"  # set this to the name of the 
# directory containing "HockeyLeague.xlsx"
file_name<-"HockeyLeague.xlsx"  # set the file name
file_path<-paste(folder_path,file_name,sep="") 
# create the file_path
loss_data_frame<-read_excel(file_path,sheet="Losses") # read of a sheet from an xl file
```
```{r}
losses_tidy<- loss_data_frame%>%
  pivot_longer(!1,names_to = 'Year',values_to = "value")%>%
  separate(value,into= c("Losses","Total"),sep="of")%>%
  rename("Team"=1)%>%
  mutate(Year = as.integer(Year),Losses = as.integer(Losses),Total = as.integer(Total))

wins_tidy%>%dim()
```

```{r}
losses_tidy%>%head(5)
```
#### Q3

```{r}

# have questions about how to use across

hockey_df<-inner_join(wins_tidy,losses_tidy)

hockey_df_new<-hockey_df%>%
  mutate(Draws= Total-(Wins +Losses),Wins_rt = round(Wins/Total,2),Losses_rt = round(Losses/Total,2), Draws_rt = round(Draws/Total,2))
  


```
#### Q4

```{r}
#have questions about how to use across

#hockey_df_new%>%
  #group_by(Team)%>%
  #select(Team,Wins_rt,Losses_rt,Draws_rt)%>%
  #summarise(across(everything()),list(mean=mean,median=median),names=c(W_mn,W_mn,L_mn,L_md,D_mn,D_md))

```

```{r}

hockey_df_new%>%
  group_by(Team)%>%
  summarise(W_md= median(Wins_rt),W_mn = round(mean(Wins_rt),3),L_md = median(Losses_rt), L_mn =round(mean(Losses_rt),3),D_md= median(Draws_rt),D_mn = round(mean(Draws_rt),3))%>%
  arrange(desc(W_md))
```

## 3 Visualisation
#### Q1

```{r}

univar_plot<- ggplot(data=filter(wins_tidy,Team =="Ducks"),aes(x=Wins)) + xlab("wins") 
univar_plot +geom_histogram(binwidth = 3) + ylab("Count")

```
#### Q2
```{r}
univar_plot+geom_density(adjust = 0.5) +ylab('Density')

```
```{r}
univar_plot+geom_density(adjust = 2) +ylab('Density')
```
####Q3


```{r}


wins_team<-wins_tidy%>%
  select(Team,Year,Wins)%>%
  pivot_wider(names_from = Team,values_from = Wins)
wins_team
```
```{r}

scatter_wins<-ggplot(data = wins_team, aes(y=Eagles,x=Ducks))+ xlab("Ducks")+ylab("Eagles")

scatter_wins + geom_point(size =3)
```

