---
title: "Assginment02"
author: "Zhen Liu"
date: "2022-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data Wrangling 
### 1.1 Select and Filter 
#### Q1
```{r load-packages, include=FALSE}
library(dplyr)
library(magrittr)
library(knitr)
library(tidyverse)
library(Stat2Data)
data("Hawks")
```


```{r}
hSF<-Hawks%>%
  filter(Species == 'RT'& Weight > 1)%>%
  select(Wing,Weight,Tail)
head(hSF)
```
#### Q2 
#### How many variables does the data frame hSF have? How mant examples does the data frame hSF have? How many observations? How many cases?

```{r}

dim(hSF)
```

totally 3 variables Wing, Weight, Tail\
totally 573 obervations per variable\
totally 1716 cases or examples

### 1.2 The arrange function

#### Q1

```{r}
head(hSF %>%  arrange((Wing)),5)

```
### 1.3 Join and rename functions 

#### Q1

```{r}
Species <- c("CH","RT","SS")
species_name_full <- c("Cooper's","Red-tailed","Sharp-shinned")
hawkSpeciesNameCodes<- data.frame(Species, species_name_full)
```

#### Q2
```{r}
hawksFullName <- left_join(Hawks,hawkSpeciesNameCodes,by = "Species")

hawksFullName<- (hawksFullName%>%select(-Species)%>%rename(Species =species_name_full ))

```
#### Q3
```{r}
head(hawksFullName%>%select(Species,Wing,Weight),7)
```
### 1.4 The mutate function
#### Q1 
\[
\Bird-BMI: = 1000 × Weight/Wing-span^2.
\]

```{r}
hawksWithBMI<- Hawks %>%
  mutate(bird_BMI = 1000*(Weight/(Wing*Wing)) )%>%
  select(Species,bird_BMI)%>%
  arrange(desc(bird_BMI))

head(hawksWithBMI,8)
```

### 1.5 Summarize and group-by functions
#### Q1 


```{r}
hawksFullName%>%
  group_by(Species)%>%
  summarise(num_rows = n(), mn_wing= mean(Wing),md_wing = median(Wing), t_mn_wing = mean(Wing,trim = 0.1),b_wt_ratio= max(Wing/Tail))



```
#### Q2

```{r}

hawksFullName%>%
  group_by(Species)%>%
  select(Wing,Weight,Culmen,Hallux,Tail,StandardTail,Tarsus,Crop)%>%
  summarise(across(everything(),~sum(is.na(.x))))


```

