---
title: "Assignment 07"
author: "Zhen Liu"
date: "2022-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Maximum likelihood estimates

### Maximum likelihood estimates for Red tailed hawks 

```{r}
library(Stat2Data)
data("Hawks")


```

#### Q1


```{r}

RedTailedDf<-Hawks%>%
  filter(Species == 'RT')%>%
  select(Weight,Tail,Wing)


head(RedTailedDf,5)

```
```{r}
dim(RedTailedDf)
```


#### Q2


```{r}

# estimate of mu and sigma

mu<-mean(RedTailedDf$Tail)
sigma<-sd(RedTailedDf$Tail)*sqrt(576/577)


mu
sigma
```

#### Q3


```{r}

weights<- seq(mu-3*sigma,mu+3*sigma,sigma*0.001)

colors<- c("MLE density" = "red","Kernel density"="blue")

estimated_density = data.frame(
  Weight = weights,
  Density = dnorm(weights, mean = mu,sd = sigma)
)


plot_obj<- ggplot() + geom_line(data =estimated_density,
                                aes(x = Weight, y= Density, color = "MLE density"))


plot_obj + geom_density(data = RedTailedDf,
                        aes(x =RedTailedDf$Tail, color = "Kernel density")) + labs(y="Density function",color = "Estimator") + theme_bw() + scale_color_manual(values = colors)

```

### Unbiased estimation of the population variance 

#### Q1

```{r}


num_trials<-1000 # set the number of trials
set.seed(0) # set the random seed
sampling_simulation<-data.frame(trial=1:num_trials) %>%
  mutate(sample_size = map(.x=trial, ~sample(c(5,100),5, replace = TRUE)))



```


#### Q2



```{r}

```


### Maximum likelihood estimation with the Possion distribution


#### Q1

The natural log likelihood function:

$$logl(\lambda) = -n\lambda + ln(\lambda)\sum_{i = 1}^{n} X_i-\sum_{i = 1}^{n}ln(X_i!)$$
derivative natural log likelihood function 

$$\frac{\partial}{\partial \lambda}logl(\lambda) =-n + \frac{1}{\lambda}\sum_{i = 1}^{n}X_i = -n+ \frac{n}{\lambda}\overline{X}$$


#### Q2

$$-n+ \frac{n}{\lambda}\overline{X} = 0 \\
\lambda = \overline{X}$$


#### Q3

```{r}

set.seed(0)

poisson_data <- data.frame('data' = rpois(1000, 0.5))

poisson_data %>% ggplot() + 
  geom_histogram(aes(x = data, 
                     y = stat(count / sum(count))), 
                     color = 'black',
                 binwidth = 1) +
  geom_vline(xintercept = 0.5, 
             size = 1, 
             linetype = 'dashed',
             color = 'red') +
  theme_bw() +
  labs(x = 'Number of successes per period',
       y = 'Proportion',
       title = '1,000 samples of Pois(lambda = 0.5)')


```


#### Q4


```{r}


Von_df<-read.csv("VonBortkiewicz.csv",header = TRUE,sep =",")

head(Von_df)
```
```{r}
dim(Von_df)
mean(Von_df$fatalities)
```

```{r}
data<-Von_df$fatalities
dpois(data[1],lambda = 1)
```

```{r}
lambda_values <- seq(-5, 5,by = 0.1)
likelihood <- dpois(data[1], lambda = lambda_values)

# arranged into a data frame
lh_single <- data.frame(lambda_values = lambda_values, likelihood = likelihood )

lh_single %>% 
  ggplot(aes(x = lambda_values,y = likelihood))+
  geom_point(size = 1, color = "blue")+
  labs(title = "Likelihood of a single data point over multiple lambda values",
       subtitle = "(we will look for a maximum)",
       caption = "Math 32",
       x = "lambda",
       y = "likelihood") +
  theme_bw()
```

### Maximum likehood estimation for the exponential distribution


#### Q1

$$\lambda_0 = \frac{1}{\overline{X}}$$

#### Q2

```{r}

CP_df<-read.csv("CustomerPurchase.csv",header = TRUE,sep =",")

dim(CP_df)

```
```{r}
time_d<-data.frame(CP_df$Time)
time_d
```

```{r}

for (i in 1:nrow(time_d)){
  time_d[i,] <- time_d[i+1,]-time_d[i,]
}
```

```{r}
time_d = as.numeric(as.character(time_d$CP_df.Time)) 
```
```{r}

CP_df%>%
  mutate(time_diffs = lead(time_d))


```

#### Q3



```{r}

x<-CP_df$Purchase

nloglik<- function(x,theta) sum(-dexp(x=x,rate=theta,log=T))

optimize(f=nloglik,x=x,interval = c(0,1000))$minimum


```

```{r}
1/mean(x)
```


#### Q4




## Confidence intervals 

### Student's t-confidence intervals


#### Q1

does not change
wider
narrower



#### Q2


```{r}
x<-c(RedTailedDf$Weight)

m<-mean(x,na.rm = TRUE)
s<-sd(x,na.rm = TRUE)
l<-length(x)
```
```{r}

error<- qt(0.99,df= l-1)*s/sqrt(l)

left<-m-error
right<- m+error
left
right
```


#### Q3

```{r}


ggplot(data = RedTailedDf,aes(x=x)) + geom_density() + theme_bw()


```
```{r}
ggplot(data = RedTailedDf,aes(sample=x)) +theme_bw() +stat_qq()
+ stat_qq_line(color = "blue")

```




### Investigating coverage for Student's intervals

#### Q1 


```{r}
student_t_confidence_interval<-function(sample,confidence_level){ 
  sample<-sample[!is.na(sample)] #removeanymissingvalues 
  n<-length(sample) #computesamplesize 
  mu_est<-mean(sample) #computesamplemean 
  sig_est<-sd(sample) #computesamplesd 
  alpha=1-confidence_level #alphafromgamma 
  t<-qt(1-alpha/2,df=n-1) #getstudenttquantile 
  l=mu_est-(t/sqrt(n))*sig_est #lower 
  u=mu_est+(t/sqrt(n))*sig_est #upper 
  return(c(l,u)) }
```


```{r}


num_trials<-100000 
sample_size<-30 
mu_0<-1 
sigma_0<-3 
alpha<-0.05 
set.seed(0) #setrandomseedforreproducibility


single_alpha_coverage_simulation_df<-data.frame(trial=seq(num_trials))%>%
  #generaterandomGaussiansamples:
  mutate(sample=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0)))%>% 
  #generateconfidenceintervals:
  mutate(ci_interval=map(.x=sample,.f=~student_t_confidence_interval(.x,1-alpha)))%>%
  #checkifintervalcoversmu_0:
  mutate(cover=map_lgl(.x=ci_interval,.f=~((min(.x)<=mu_0)&(max(.x)>=mu_0))))%>% 
  #computeintervallength:
  mutate(ci_length=map_dbl(.x=ci_interval,.f=~(max(.x)-min(.x))))



#estimateofcoverageprobability:
single_alpha_coverage_simulation_df%>% 
  pull(cover)%>%
  mean()


```




## One sample hypothesis testing

### One sample t-test on penguins data

#### Q1 

```{r}


library(palmerpenguins)

data(package = 'palmerpenguins')

head(penguins)

```


```{r}
bill_adelie<-penguins%>%
  filter(species == 'Adelie')%>%
  select(bill_length_mm)
bill_adelie<-bill_adelie$bill_length_mm

```



```{r}

sample_size<-length(bill_adelie)

sample_mean<-mean(bill_adelie,na.rm = TRUE)

sample_sd<-sd(bill_adelie,na.rm = TRUE)

test_statistic<-(sample_mean-40)/(sample_sd/sqrt(sample_size))


test_statistic
```

```{r}
# compute the p-value
2*(1-pt(abs(test_statistic),df = sample_size-1))
```




```{r}
t.test(x=bill_adelie,mu = 40)
```

```{r}

alpha<-0.01
t<-qt(1-alpha/2,df = sample_size-1)
confidence_interval_l<-sample_mean-t*sample_sd/sqrt(sample_size)
confidence_interval_u<-sample_mean+t*sample_sd/sqrt(sample_size)
confidence_interval<-c(confidence_interval_l,confidence_interval_u)


confidence_interval


```

### Implementing a one-sample t-test


#### Q1

```{r}

t.test(x=bill_adelie,mu =40,alternative = "two.sided")


```




