---
title: "Assginment03"
author: "Zhen Liu"
date: "2022-10-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Exploratory data analysis




```{r}

library(tidyverse)
library(Stat2Data)
data("Hawks")
```

```{r}
head(Hawks)

```
### 1.1 Location estimators

#### Q1 

```{r}


HawksTail<-Hawks[['Tail']]

HawksTail
```
```{r}

class(HawksTail)

mean(HawksTail)

median(HawksTail)
```
### 1.2 Combining location estimators with the summarise function 
#### Q1

```{r}

Hawks%>%
  summarise(Wing_mean= mean(Wing,na.rm = TRUE),Wing_t_mean = mean(Wing,na.rm = TRUE,trim = 0.5), Wing_med = median(Wing,na.rm = TRUE),Weight_mean= mean(Weight,na.rm = TRUE),Weight_t_mean = mean(Weight,na.rm = TRUE,trim = 0.5), Weight_med = median(Weight,na.rm = TRUE))


```
#### Q2
```{r}


Hawks%>%
  group_by(Species)%>%
  summarise(Wing_mean= mean(Wing,na.rm = TRUE),Wing_t_mean = mean(Wing,na.rm = TRUE,trim = 0.5), Wing_med = median(Wing,na.rm = TRUE),Weight_mean= mean(Weight,na.rm = TRUE),Weight_t_mean = mean(Weight,na.rm = TRUE,trim = 0.5), Weight_med = median(Weight,na.rm = TRUE))



```

### 1.3 Location and dispersion estimators under linear transformations 

#### Q1

```{r}

a<- 2
b<-3

mean(HawksTail*a +b)

mean(HawksTail)

```
The mean of HawksTail*a +b is almost double times as HawksTail

#### Q2

```{r}

var(HawksTail*a +b)
var(HawksTail)

sd(HawksTail*a +b)
sd(HawksTail)
```
The variance of HawksTail*a +b is almost four times as HawksTail
The standard deviation of HawksTail*a +b is almost two times as HawksTail

### 1.4 Robustness of location estimators


```{r}

hal<-Hawks$Hallux # Extract the vector of hallux lengths 
hal<-hal[!is.na(hal)] # Remove any nans

```

```{r}
outlier_val<-100 
num_outliers<-10

corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))

mean(hal)
mean(corrupted_hal)


```

```{r}
num_outliers_vect <- seq(0,1000) 
means_vect <- c() 
for(num_outliers in num_outliers_vect){
  corrupted_hal <-c(hal,rep(outlier_val,times=num_outliers)) 
  means_vect <- c(means_vect, mean(corrupted_hal))
}
```

#### Q1

```{r}

medians_vect<-c()
for(num_outliers in num_outliers_vect){
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  medians_vect<-c(medians_vect,median(corrupted_hal))
}

```

#### Q2

```{r}
t_means_vect<-c()
for(num_outliers in num_outliers_vect){
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  t_means_vect<-c(t_means_vect,mean(corrupted_hal,trim = 0.1))
}
```

#### Q3
```{r}

df_means_medians <-
  data.frame(num_outliers=num_outliers_vect,
             mean=means_vect, t_mean=t_means_vect,
             median=medians_vect)


```

```{r}

df_means_medians %>% pivot_longer(!num_outliers,
                                  names_to = "Estimator", values_to = "Value") %>%
  ggplot(aes(x=num_outliers,color=Estimator,
             linetype=Estimator,y=Value)) +
  geom_line()+xlab("Number of outliers")



```
### 1.5 Box plots and outliers

```{r}


ggplot(Hawks,aes(x=Species,y=Weight)) +
  geom_boxplot()+xlab("Species") + ylab("Weight")



```
#### Q2  quantile annd boxplots

```{r}

q =c(0.25,0.5,0.75)

Hawks%>%
  group_by(Species)%>%
  summarise(quantile025 = quantile(Weight, probs = q[1],na.rm=TRUE),
            quantile050 = quantile(Weight, probs = q[2],na.rm=TRUE),
            quantile075 = quantile(Weight, probs = q[3],na.rm=TRUE))

```
#### Q3  outliers

```{r}
num_outliers<- function(input_sample){
  quantile25<- quantile(input_sample,0.25, na.rm = TRUE)
  quantile75<- quantile(input_sample,0.75, na.rm = TRUE)
  iq_range<- quantile75 - quantile25
  outliers <- input_sample[(input_sample>quantile75+1.5*iq_range) | (input_sample<quantile25-1.5*iq_range)]
  n<-length(which(!is.na(outliers)))
  return(n)
}

num_outliers(c(0,40,60,185))
```
#### Q4 Outliers by group 


```{r}
Hawks%>%
  group_by(Species)%>%
  summarise(num_outliers_weight=num_outliers(Weight))


```

### 1.6 Covariance and correlation

#### Q1

```{r}

cov(Hawks$Weight,Hawks$Wing,use = 'complete.obs')

cor(Hawks$Weight,Hawks$Wing,use = 'complete.obs')

```
#### Q2

```{r}
S_new_n<- function(X,Y,a,b,c,d){
  S <- cov(X,Y,use ='complete.obs' )
  x_new<- a*X+b
  y_new<- c*Y+d
  S_new <- abs(cov(x_new,y_new,use ='complete.obs'))
  
  return(S_new)
}
```
```{r}

S_new_n(Hawks$Weight,Hawks$Wing,2.4,7.1,-1,3)

```

S_new is 2.4 times than S

## 2. Random experiments, events and sample spaces, and the set theory

### 2.1 Random experiments, events and sample sapces

#### Q1 Firstly, write down the definition of a random experiment, event and sample space.

A random experiment is a procedure (real or imagined) which: 
1. has a well-defined set of possible outcomes; 
2. could (at least in principle) be repeated arbitrarily many times.

An event is a set (i.e. a collection) of possible outcomes of an experiment

A sample space is the set of all possible outcomes of interest for a random experiment


#### Q2 Consider a random experiment of rolling a dice twice. Give an example of what is an event in this random experiment. Also, can you write down the sample space as a set? What is the total number of different events in this experiment? Is the empty set considered as an event?

Example event : get {6,6}
Total number of different events : 36
you cannot get an empty set because you must get a number when you rolling a dice 

### 2.2 Set theory

#### Q1 Set operations:

Let the sets A, B, C be defined by A := {1, 2, 3}, B := {2, 4, 6}, C := {4, 5, 6}. 

1. What are the unions A∪B and A∪C? 
2. What are the intersections A ∩ B and A∩C? 
3. What are the complements A\B and A\C? 
4. Are A and B disjoint? Are A and C disjoint? 
5. Are B and A\B disjoint? 
6. Write down a partition of {1,2,3,4,5,6} consisting of two sets. Also, write down another partition of {1,2,3,4,5,6} consisting of three sets


Answer: 
1. unions A∪B : {1,2,3,4,6}    A∪C :{1,2,3,4,5,6}
2. intersections  A ∩ B : {2}    A∩C  empty set 
3. complements A\B : {1,3}    A\C: {1,2,3}
4. no  / yes
5. yes
6. {1,2,3}   {4,5,6}
{1,2} {3,4} {5,6}


#### Q2 Complements, subsets and De Morgan’s laws

Let Ω be a sample space. Recall that for an event A ⊆ Ω the complement Ac := Ω\A := {w ∈ Ω : w /∈ A}. Take a pair of events A ⊆ Ω and B ⊆ Ω.
1. Can you give an expression for (Ac)c without using the notion of a complement? 
2. What is Ωc? 
3. (Subsets) Show that if A ⊆ B, then Bc ⊆ Ac. 
4. (De Morgan’s laws) Show that (A ∩ B)c = Ac ∪ Bc. Let’s suppose we have a sequence of events A1,A2,··· ,AK ⊂ Ω. Can you write out an expression for (∩K k=1Ak)c? 
5. (De Morgan’s laws) Show that (A∪B)c = Ac ∩Bc. 
6. Let’s suppose we have a sequence of events A1,A2,··· ,AK ⊂ Ω. Can you write out an expression for (∪K k=1Ak)c?


Answer:
1. A
2. empty set
3. A: {1,2,3} B{1,2,3,4,5} Ω:{1,2,3,4,5,6,7,8,9}
   Ac :{4,5,6,7,8,9}   Bc:{6,7,8,9}
   Bc ⊆ Ac
4. $$(∩_{k=1}^{K} A_k)^c := ∪_{k=1}^{K} (A_k)^c$$

5. $$A∪B: = \{1,2,3,4,5\}$$ 
$$(A∪B)^c: = \{6,7,8,9\}$$ 
$$A^c: = \{4,5,6,7,8,9\}\     
B^c : =\{6,7,8,9\} \
A^c ∩ B^c := \{6,7,8,9\}
$$ 
6. $$(∪_{k=1}^{K} A_k)^c := ∩_{k=1}^{K} (A_k)^c$$

#### Q3 Cardinality and the set of all subsets:


$$ E := \{A∈Ω :A∈A_K &nbsp; &nbsp; for all i \}$$


