---
title: "Assignment 08"
author: "Zhen Liu"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


## Obstacles to valid scientific inference

### Q1


1. Measurement distortions

your data has errors

Measurement error is very common in practice
- Miscalibration of measurement instruments e.g. a clock running slightly too fast or slow.
- Rounding errors due to computational constraints e.g. 0.904 converted to 0.9.
- Inaccurate responses to questionnaires e.g. under reporting alcohol consumption.
- Human error e.g. data entry mistakes

2. Selection bias

Selection bias occurs when the data included in the analysis misrepresents the underlying population of interest.
sample bias, self-selection bias, attrition bias, and post-hoc selection

sample bias: You want to know what the most popular genre (jazz, folk, rock …) is amongst attendees of a music festival

self selection bias: online reviews of restaurants might disproportionately represent subsets of the population with strong opinions or certain age groups

attrition bias:A scientist is investigating the efficacy of a new exercise program

post hoc selection: A scientist is investigating the efficacy of medical treatment


Randomized samples: sampling from the population of interest which is the set of all attendees of a music festival


3. Confounding variables

Correlation does not imply causation

Example: the increased sales of sun glass may not be the reason for the increased sales of ice cream, but the sunny weather is


## paired t-test and effect size

```{r}

library(PairedData)
data("Barley")
detach('package:PairedData',unload = TRUE)
detach('package:MASS', unload = TRUE)

head(Barley,4)

```

### Q1

```{r}
t.test(x =Barley$Glabron,y =Barley$Velvet,paired = TRUE )
```
The p-value is 0.0118 which is above the significant level of 0.01. Thus we do not reject the null hypothesis

### Q2




